Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the `eechidna` `R` package.

### 2019-03-29 draft-paper.Rmd
new EBS working paper files

--- 
title: "Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the `eechidna` `R` package."
author: "Jeremy Forbes, Di Cook & Rob Hyndman"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
classoption: openany
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is an incomplete draft of a paper to submitted to ANZJS."
---

<!--chapter:end:index.Rmd-->

@@

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center', cache = T, message = F)
library(tidyverse)
library(eechidna)
library(ggplot2)
library(ggthemes)
library(knitr)
library(nlme)
library(rgeos)
library(spdep)

@@

```{r wrangle, include = F, message = F}
# Function to standardize variables
standardise_vars <- function(df) {
  hold <- df
  num_cols <- sapply(df, class) == 'numeric'
  df[, num_cols] <- lapply(df[, num_cols], scale)
  names(df) <- names(hold)
  df$LNP_Percent <- hold$LNP_Percent
  return(df)
}

@@

<!--chapter:end:01-intro.Rmd-->

# Data collection {#data}

## Voting data

@@

```{r timeline, message = F, echo = F, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
timeline <- data.frame(year = c(2001:2016)) %>%
  mutate(Election = ifelse(year %in% c(2001, 2004, 2007, 2010, 2013, 2016),1,0),
         Census = ifelse(year %in% c(2001, 2006, 2011, 2016), 1.4, 0)) %>%
  gather(c(Election,Census), key = event, value = measure) %>%  filter(measure != 0) %>%
  mutate(year_vis = ifelse(event == "Census", year + 0.1, year))

@@

<!--chapter:end:03-imputation.Rmd-->

---
output:
  pdf_document: default
  html_document: default
---
# Modelling {#modelling}

@@

Table 4.1 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.

```{r printmodels, eval = F}
# Add rho and make adjustments to this
stargazer::stargazer(glsmod16, glsmod13, glsmod10, glsmod07, glsmod04, glsmod01, header = FALSE, 
  column.sep.width = "1pt", 
  title = "Estimated model for each of the six elections.",
  dep.var.labels = "Two-party preferred vote in favor of the Liberal party", 
  font.size = "scriptsize", digits = 2, 
  column.labels = c("2001", "2004", "2007", "2010", "2013", "2016"),
  notes = c("Estimated coefficients for variable named in column one shown for", "election year indicated by column heading, with estimated standard", "deviation for each coefficient shown below in parenthesis. Overall", "summary measures for each regression equation are provided in the", "bottom panel."),
  notes.align = "l")
```

\begin{table}[!htbp] \centering 
  \caption{Estimated model for each of the six elections.} 
  \label{} 
\tiny 
\begin{tabular}{@{\extracolsep{1pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{6}{c}{Two-party preferred vote in favor of the Liberal party} \\ 
 & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
\hline \\[-1.8ex] 
 $\rho$ & 0.48$^{***}$ & 0.29$^{*}$ & 0.15 & $-$0.04 & 0.15 & 0.44$^{***}$ \\ 
  & (0.16) & (0.16) & (0.17) & (0.19) & (0.17) & (0.15) \\ 
  & & & & & & \\ 
 AusCitizen & $-$1.53 & $-$0.87 & 1.00 & 0.67 & $-$2.83 & $-$2.77 \\ 
  & (2.95) & (2.90) & (2.93) & (2.57) & (2.63) & (2.51) \\ 
  & & & & & & \\ 
 Born\_Asia & $-$2.13 & $-$2.56 & $-$7.58$^{***}$ & $-$1.37 & 0.74 & 2.63 \\ 
  & (2.20) & (2.58) & (2.68) & (1.96) & (2.55) & (3.03) \\ 
  & & & & & & \\ 
 Born\_MidEast & $-$0.82 & 1.70 & 1.79 & 2.92$^{*}$ & 3.22$^{*}$ & 1.92 \\ 
  & (1.56) & (1.54) & (1.54) & (1.51) & (1.78) & (1.53) \\ 
  & & & & & & \\ 
 Born\_SE\_Europe & $-$1.63 & $-$1.25 & $-$1.72 & $-$0.20 & 0.04 & 1.02 \\ 
  & (0.99) & (1.15) & (1.34) & (1.17) & (2.05) & (2.34) \\ 
  & & & & & & \\ 
 Born\_UK & $-$1.37 & $-$0.35 & $-$0.29 & 0.29 & $-$0.87 & $-$0.49 \\ 
  & (0.98) & (1.00) & (1.05) & (0.80) & (0.90) & (1.04) \\ 
  & & & & & & \\ 
 BornElsewhere & 1.74 & 4.04 & 9.78$^{**}$ & 5.01 & $-$0.51 & $-$1.08 \\ 
  & (4.52) & (4.46) & (4.47) & (3.59) & (3.67) & (3.40) \\ 
  & & & & & & \\ 
 Buddhism & 1.51 & 1.09 & 2.41$^{*}$ & 0.90 & 2.84$^{*}$ & 1.50 \\ 
  & (1.24) & (1.25) & (1.40) & (1.25) & (1.61) & (1.94) \\ 
  & & & & & & \\ 
 Christianity & 5.17$^{**}$ & 3.08 & $-$0.60 & $-$1.93 & 3.44 & 0.53 \\ 
  & (2.52) & (2.79) & (3.44) & (3.45) & (4.19) & (4.58) \\ 
  & & & & & & \\ 
 CurrentlyStudying & $-$0.09 & 1.73 & 3.28$^{***}$ & 4.03$^{***}$ & 1.69 & $-$1.14 \\ 
  & (1.13) & (1.17) & (1.21) & (1.12) & (1.09) & (1.01) \\ 
  & & & & & & \\ 
 DeFacto & $-$11.02$^{***}$ & $-$9.73$^{***}$ & $-$8.60$^{***}$ & $-$6.55$^{***}$ & $-$5.04$^{**}$ & $-$6.42$^{***}$ \\ 
  & (2.55) & (2.62) & (2.57) & (1.90) & (2.00) & (1.69) \\ 
  & & & & & & \\ 
 DiffAddress & 6.93$^{***}$ & 5.02$^{***}$ & 4.72$^{***}$ & 3.40$^{***}$ & 4.63$^{***}$ & 3.86$^{***}$ \\ 
  & (1.50) & (1.71) & (1.60) & (0.88) & (1.03) & (0.95) \\ 
  & & & & & & \\ 
 Distributive & 1.65 & 2.76$^{**}$ & 2.88$^{**}$ & 2.25$^{**}$ & 2.75$^{**}$ & 1.67 \\ 
  & (1.17) & (1.25) & (1.30) & (1.02) & (1.13) & (1.08) \\ 
  & & & & & & \\ 
 Education & $-$3.04 & $-$0.90 & $-$1.82 & $-$0.77 & 1.94 & 0.62 \\ 
  & (2.99) & (3.04) & (3.35) & (2.89) & (2.86) & (2.23) \\ 
  & & & & & & \\ 
 Extractive & 6.81$^{***}$ & 6.09$^{***}$ & 6.97$^{***}$ & 5.81$^{***}$ & 5.31$^{***}$ & 4.69$^{***}$ \\ 
  & (1.31) & (1.42) & (1.48) & (1.16) & (1.23) & (1.36) \\ 
  & & & & & & \\ 
 FamHouseSize & $-$4.35 & $-$6.16$^{*}$ & $-$8.22$^{**}$ & $-$9.67$^{***}$ & $-$5.11$^{*}$ & $-$2.94 \\ 
  & (2.86) & (3.15) & (3.28) & (2.66) & (2.78) & (2.27) \\ 
  & & & & & & \\ 
 Incomes & 8.46$^{***}$ & 4.19 & 2.73 & 4.36$^{*}$ & 2.95 & 2.99$^{*}$ \\ 
  & (2.57) & (2.68) & (3.06) & (2.56) & (2.60) & (1.76) \\ 
  & & & & & & \\ 
 Indigenous & 4.47$^{***}$ & 6.12$^{***}$ & 5.90$^{***}$ & 6.81$^{***}$ & 5.62$^{***}$ & 3.97$^{***}$ \\ 
  & (1.38) & (1.49) & (1.49) & (1.17) & (1.35) & (1.26) \\ 
  & & & & & & \\ 
 Islam & 2.64 & 1.20 & $-$1.04 & $-$1.21 & 0.61 & 0.15 \\ 
  & (1.72) & (1.71) & (1.94) & (1.80) & (2.00) & (2.03) \\ 
  & & & & & & \\ 
 Judaism & 1.96$^{***}$ & 1.60$^{*}$ & 0.46 & 0.01 & 1.74 & 1.02 \\ 
  & (0.71) & (0.85) & (1.05) & (1.00) & (1.26) & (1.38) \\ 
  & & & & & & \\ 
 ManagerAdminClericalSales & 4.62$^{***}$ & 3.99$^{***}$ & 4.32$^{***}$ & 5.21$^{***}$ & 3.28$^{***}$ & 2.10$^{***}$ \\ 
  & (1.03) & (0.99) & (1.00) & (0.78) & (0.85) & (0.70) \\ 
  & & & & & & \\ 
 Married & $-$1.39 & 1.46 & 5.07$^{*}$ & 5.06$^{*}$ & 5.50$^{**}$ & 2.27 \\ 
  & (2.64) & (2.71) & (2.91) & (2.59) & (2.79) & (2.26) \\ 
  & & & & & & \\ 
 MedianAge & 2.94$^{*}$ & 2.28 & 2.08 & 1.86 & 3.73$^{**}$ & 1.54 \\ 
  & (1.63) & (1.88) & (2.08) & (1.65) & (1.55) & (1.30) \\ 
  & & & & & & \\ 
 NoReligion & 5.76$^{**}$ & 3.53 & 0.25 & $-$0.31 & 3.43 & 1.06 \\ 
  & (2.59) & (2.58) & (2.98) & (2.90) & (3.45) & (3.85) \\ 
  & & & & & & \\ 
 OneParent\_House & $-$0.85 & 0.11 & 0.22 & 1.66 & 1.37 & $-$1.00 \\ 
  & (1.39) & (1.46) & (1.61) & (1.38) & (1.49) & (1.34) \\ 
  & & & & & & \\ 
 OtherLanguageHome & 1.27 & $-$4.69 & $-$4.37 & $-$6.02 & $-$6.76 & $-$10.40$^{**}$ \\ 
  & (4.95) & (5.50) & (5.70) & (4.44) & (5.61) & (4.97) \\ 
  & & & & & & \\ 
 PropertyOwned & 1.76 & 0.68 & $-$1.17 & 0.78 & $-$0.92 & $-$0.76 \\ 
  & (1.38) & (1.43) & (1.49) & (1.09) & (1.21) & (1.19) \\ 
  & & & & & & \\ 
 RentLoanPrice & $-$0.70 & 1.01 & 2.12 & $-$0.54 & $-$1.54 & $-$0.31 \\ 
  & (2.02) & (1.96) & (1.96) & (1.48) & (1.68) & (1.50) \\ 
  & & & & & & \\ 
 SocialServ & 2.90$^{***}$ & 0.94 & 1.91 & 1.77 & 1.89 & 2.42$^{*}$ \\ 
  & (1.10) & (1.22) & (1.35) & (1.14) & (1.27) & (1.27) \\ 
  & & & & & & \\ 
 Transformative & 4.75$^{***}$ & 3.74$^{**}$ & 3.28$^{*}$ & 5.20$^{***}$ & 5.74$^{***}$ & 2.99$^{**}$ \\ 
  & (1.41) & (1.57) & (1.74) & (1.47) & (1.59) & (1.41) \\ 
  & & & & & & \\ 
 Unemployment & 0.36 & $-$1.22 & $-$2.42 & $-$2.88$^{**}$ & $-$5.09$^{***}$ & $-$3.00$^{**}$ \\ 
  & (1.42) & (1.54) & (1.73) & (1.41) & (1.51) & (1.28) \\ 
  & & & & & & \\ 
 Constant & 50.45$^{***}$ & 53.48$^{***}$ & 49.92$^{***}$ & 47.32$^{***}$ & 52.63$^{***}$ & 50.78$^{***}$ \\ 
  & (0.74) & (0.56) & (0.51) & (0.35) & (0.45) & (0.67) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
Observations & 150 & 150 & 150 & 150 & 150 & 150 \\ 
Log Likelihood & $-$399.52 & $-$405.41 & $-$418.08 & $-$398.73 & $-$403.34 & $-$398.33 \\ 
Akaike Inf. Crit. & 863.03 & 874.83 & 900.16 & 861.46 & 870.68 & 860.67 \\ 
Bayesian Inf. Crit. & 951.96 & 963.76 & 989.10 & 950.40 & 959.61 & 949.60 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{6}{l}{Estimated coefficients for variable named in column one shown for} \\ 
 & \multicolumn{6}{l}{election year indicated by column heading, with estimated standard} \\ 
 & \multicolumn{6}{l}{deviation for each coefficient shown below in parenthesis. Overall} \\ 
 & \multicolumn{6}{l}{summary measures for each regression equation are provided in the} \\ 
 & \multicolumn{6}{l}{bottom panel.} \\ 
\end{tabular} 
\end{table}

<!--chapter:end:04-modelling.Rmd-->

# Results
```{r rhovis}
rho_df <- bind_rows(glsmod16$rho_df %>% mutate(year = "2016"), 
  glsmod13$rho_df %>% mutate(year = "2013"),
  glsmod10$rho_df %>% mutate(year = "2010"), 
  glsmod07$rho_df %>% mutate(year = "2007"),
  glsmod04$rho_df %>% mutate(year = "2004"), 
  glsmod01$rho_df %>% mutate(year = "2001")) %>% 
  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se)

rho_df %>% ggplot() + 
  geom_point(aes(x = year, y = estimate, col = factor(p < 0.05)), size = 3) +
  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95, col = factor(p < 0.05)), size = 1.5) + 
  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) +
  scale_color_manual(values = c("grey50", "black")) +
  guides(col = F)
```

The spatial autoregressive coefficient $\rho$ is significant in only the 2001 and 2016 elections, meaning that in these years an electorate's political climate was affected by the attitudes of it's neighbours. Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no spatial effect, meaning electorates effectively acted independently.

To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used. These show the direction, size and significance of an estimated effect, as the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band. Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs. Although topical in the lead up to an election, incomes only significantly affected electoral two-party preference in 2016 (see Figure \@ref(fig:plotincomes)), with higher income areas associated with support for the Liberal party. In the prior elections, the estimated effect is still positive, albeit insignificant. Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections(Figure \@ref(fig:plotunemploy)), with electorates with higher unemployment favouring Labor. This effect has become weaker over time, with unemployment having an estimated effect of virtually zero in 2016.

```{r myvisreg, include = F}
# Function to produce visreg style conditional plots
my_visreg <- function(my_model, sp_weights, varname, 
  nolabs = FALSE, xlimits = NULL, ylimits = NULL, year = "") {

@@

  if (!is.null(xlimits) & !is.null(ylimits)) {
    myplot <- myplot + coord_cartesian(xlim = xlimits, ylim = ylimits)
  }

  return(myplot)
}

>>>>

---
title: "Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the `eechidna` `R` package."
author:
- familyname: Forbes
  othernames: Jeremy
  address: Monash University\newline Melbourne, Australia
  email: jeremyforbes1995@gmail.com
  correspondingauthor: true
- familyname: Cook
  othernames: Di
  address: Monash University\newline Melbourne, Australia
- familyname: Hyndman
  othernames: Rob
  address: Monash University\newline Melbourne, Australia
abstract: "A brief summary of our ideas"
keywords: "blah, blah"
wpnumber: no/yr
jelcodes: C10,C14,C22
blind: false
cover: true
toc: false
bibliography: references.bib
biblio-style: authoryear-comp
header-includes: \usepackage{float}
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
---

@@

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center', cache = T, message = F)
library(tidyverse)
library(eechidna)
library(ggplot2)
library(ggthemes)
library(knitr)
library(nlme)
library(rgeos)
library(spdep)
library(gridExtra)
library(grid)
library(predictmeans)

@@

```{r wrangle, include = F}
# Function to standardize variables
standardise_vars <- function(df) {
  hold <- df
  num_cols <- sapply(df, class) == 'numeric'
  df[, num_cols] <- lapply(df[, num_cols], scale)
  names(df) <- names(hold)
  df$LNP_Percent <- hold$LNP_Percent
  return(df)
}

# Take log of indigneous, judaism, islam, buddhism
abs2016 <- abs2016 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
abs2013 <- abs2013 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
abs2010 <- abs2010 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
abs2007 <- abs2007 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
abs2004 <- abs2004 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
abs2001 <- abs2001 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))

@@

# Data collection, wrangling and imputation {#data}

## Collecting the data

@@

```{r timeline, echo = F, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
timeline <- data.frame(year = c(2001:2016)) %>%
  mutate(Election = ifelse(year %in% c(2001, 2004, 2007, 2010, 2013, 2016),1,0),
         Census = ifelse(year %in% c(2001, 2006, 2011, 2016), 1.4, 0)) %>%
  gather(c(Election,Census), key = event, value = measure) %>%  filter(measure != 0) %>%
  mutate(year_vis = ifelse(event == "Census", year + 0.1, year))

@@

# Modelling {#modelling}

@@

Table 3 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.

```{r printmodels, eval = F}
# Add rho and GLS residual standard error
stargazer::stargazer(glsmod01, glsmod04, glsmod07, glsmod10, glsmod13, glsmod16, header = FALSE, 
  column.sep.width = "1pt", 
  title = "Estimated model for each of the six elections.",
  dep.var.labels = "Two-party preferred vote in favor of the Liberal party", 
  font.size = "scriptsize", digits = 2, 
  column.labels = c("2001", "2004", "2007", "2010", "2013", "2016"),
  notes = c("Estimated coefficients for variable named in column one shown for", "election year indicated by column heading, with estimated standard", "deviation for each coefficient shown below in parenthesis. Overall", "summary measures for each regression equation are provided in the", "bottom panel."),
  notes.align = "l")
```
\begin{table}[!htbp] \centering 
  \caption{Estimated model for each of the six elections.} 
  \label{} 
\scriptsize 
\begin{tabular}{@{\extracolsep{1pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{6}{c}{Two-party preferred vote in favor of the Liberal party} \\ 
 & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\ 
\hline \\[-1.8ex] 
 $\rho$ & 0.46$^{***}$ & 0.29$^{*}$ & 0.24 & 0.19 & 0.27$^{*}$ & 0.50$^{***}$ \\ 
  & (0.15) & (0.17) & (0.17) & (0.16) & (0.16) & (0.17) \\

 AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\ 
  & (2.26) & (2.43) & (2.34) & (2.79) & (2.76) & (2.71) \\

 Born\_Asia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83$^{**}$ & $-$3.03 & $-$0.55 \\ 
  & (2.18) & (2.44) & (2.19) & (2.73) & (2.71) & (2.17) \\

 Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01$^{*}$ & $-$2.03 & $-$0.92 & $-$1.44 \\ 
  & (1.07) & (1.20) & (1.11) & (1.27) & (1.24) & (1.13) \\

 Born\_SE\_Europe & $-$3.21$^{**}$ & $-$4.24$^{***}$ & $-$3.61$^{***}$ & $-$4.14$^{***}$ & $-$3.69$^{***}$ & $-$2.72$^{***}$ \\ 
  & (1.42) & (1.46) & (1.02) & (1.19) & (1.07) & (0.97) \\

 Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\ 
  & (1.00) & (0.98) & (0.90) & (1.07) & (1.04) & (1.04) \\

 BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\ 
  & (3.30) & (3.68) & (3.38) & (4.23) & (4.15) & (3.97) \\

 Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\ 
  & (1.39) & (1.61) & (1.61) & (1.80) & (1.68) & (1.56) \\

 Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\ 
  & (1.73) & (1.85) & (1.83) & (1.99) & (1.85) & (1.78) \\

 CurrentlyStudying & $-$2.19$^{**}$ & $-$0.13 & 2.06$^{*}$ & 2.12$^{*}$ & 1.15 & $-$0.16 \\ 
  & (0.99) & (1.13) & (1.17) & (1.25) & (1.26) & (1.18) \\

 DeFacto & $-$6.44$^{***}$ & $-$5.37$^{**}$ & $-$6.43$^{***}$ & $-$8.07$^{***}$ & $-$6.56$^{**}$ & $-$8.53$^{***}$ \\ 
  & (1.87) & (2.48) & (2.31) & (3.06) & (3.11) & (2.83) \\

 DiffAddress & 3.88$^{***}$ & 5.06$^{***}$ & 4.22$^{***}$ & 5.57$^{***}$ & 3.53$^{*}$ & 5.67$^{***}$ \\ 
  & (0.94) & (1.12) & (0.99) & (1.76) & (1.91) & (1.60) \\

 Distributive & 1.27 & 2.01$^{*}$ & 1.36 & 1.57 & 2.10$^{*}$ & 1.20 \\ 
  & (1.12) & (1.21) & (1.13) & (1.34) & (1.27) & (1.21) \\

 Education & 1.08 & 0.52 & $-$5.52$^{*}$ & $-$4.08 & $-$4.44 & $-$7.07$^{**}$ \\ 
  & (2.38) & (3.12) & (3.27) & (3.95) & (3.78) & (3.55) \\

 Extractive & 4.83$^{***}$ & 5.45$^{***}$ & 5.37$^{***}$ & 7.31$^{***}$ & 6.71$^{***}$ & 7.43$^{***}$ \\ 
  & (1.48) & (1.42) & (1.36) & (1.56) & (1.47) & (1.39) \\

 FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\ 
  & (2.19) & (2.72) & (2.69) & (3.25) & (3.28) & (2.90) \\

 Incomes & 4.36$^{**}$ & 5.03$^{*}$ & 9.45$^{***}$ & 7.09$^{**}$ & 7.97$^{***}$ & 12.20$^{***}$ \\ 
  & (1.77) & (2.66) & (2.75) & (3.25) & (2.92) & (2.75) \\

 Indigenous & 2.91$^{*}$ & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\ 
  & (1.68) & (1.95) & (1.75) & (2.16) & (2.14) & (2.00) \\

 Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\ 
  & (1.22) & (1.36) & (1.27) & (1.52) & (1.42) & (1.34) \\

 Judaism & 1.88$^{*}$ & 1.78 & 2.66$^{***}$ & 1.97$^{*}$ & 2.74$^{**}$ & 1.65$^{*}$ \\ 
  & (1.05) & (1.13) & (1.01) & (1.15) & (1.10) & (1.00) \\

 ManagerAdminClericalSales & 2.06$^{***}$ & 3.32$^{***}$ & 6.00$^{***}$ & 5.47$^{***}$ & 5.04$^{***}$ & 5.78$^{***}$ \\ 
  & (0.71) & (0.93) & (0.90) & (1.08) & (1.03) & (1.06) \\

 Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\ 
  & (2.31) & (2.96) & (2.83) & (3.15) & (3.03) & (2.81) \\

 MedianAge & 2.32$^{*}$ & 4.96$^{***}$ & 3.66$^{**}$ & 4.00$^{*}$ & 2.30 & 2.87 \\ 
  & (1.32) & (1.65) & (1.81) & (2.26) & (2.08) & (1.79) \\

 NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\ 
  & (1.59) & (1.71) & (1.73) & (1.92) & (1.94) & (2.04) \\

 OneParent\_House & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\ 
  & (1.36) & (1.59) & (1.49) & (1.69) & (1.57) & (1.47) \\

 OtherLanguageHome & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\ 
  & (3.22) & (4.16) & (3.91) & (4.76) & (4.66) & (4.44) \\

 PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\ 
  & (1.37) & (1.50) & (1.43) & (1.76) & (1.67) & (1.50) \\

 RentLoanPrice & $-$1.57 & $-$3.32$^{*}$ & $-$4.01$^{**}$ & $-$0.97 & $-$0.70 & $-$0.89 \\ 
  & (1.49) & (1.76) & (1.67) & (2.07) & (2.07) & (2.16) \\

 SocialServ & 2.51$^{*}$ & 1.65 & 2.47$^{*}$ & 2.53$^{*}$ & 2.35$^{*}$ & 4.45$^{***}$ \\ 
  & (1.33) & (1.41) & (1.29) & (1.47) & (1.32) & (1.19) \\

 Transformative & 3.24$^{**}$ & 4.73$^{***}$ & 4.84$^{***}$ & 4.46$^{**}$ & 3.56$^{**}$ & 4.58$^{***}$ \\ 
  & (1.55) & (1.78) & (1.74) & (1.98) & (1.78) & (1.53) \\

 Unemployment & $-$2.45$^{*}$ & $-$3.07$^{*}$ & 0.29 & 0.08 & 1.67 & 2.79$^{**}$ \\ 
  & (1.40) & (1.63) & (1.51) & (1.76) & (1.51) & (1.37) \\

 Constant & 50.81$^{***}$ & 52.60$^{***}$ & 47.31$^{***}$ & 49.93$^{***}$ & 53.51$^{***}$ & 50.49$^{***}$ \\ 
  & (0.71) & (0.58) & (0.52) & (0.57) & (0.58) & (0.80) \\

\hline \\[-1.8ex] 
Observations & 150 & 150 & 150 & 150 & 150 & 150 \\ 
Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

\newpage

# Results

## Spatial autoregressive parameter

The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections (Figure \@ref(fig:rhovis)), meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours. Conversely, in the other four elections, the spatial effect weakens to become insignficant. In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates voted effectively independently.

```{r rhovis, fig.cap="Estimates of the spatial autoregressive parameter for each of the six elections, with a 95\\% confidence interval. Only in 2001 and 2016 is there a significant spatial component", out.height='15%', out.extra = ''}
rho_df <- bind_rows(glsmod16$rho_df %>% mutate(year = "2016"), 
  glsmod13$rho_df %>% mutate(year = "2013"),
  glsmod10$rho_df %>% mutate(year = "2010"), 
  glsmod07$rho_df %>% mutate(year = "2007"),
  glsmod04$rho_df %>% mutate(year = "2004"), 
  glsmod01$rho_df %>% mutate(year = "2001")) %>% 
  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se)

rho_df %>% ggplot() + 
  geom_point(aes(x = year, y = estimate), size = 3) +
  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) + 
  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) +
  scale_color_manual(values = c("grey50", "black")) +
  labs(x = "Election year", y = "Estimate of spatial autoregressive parameter") +
  lims(y = c(-1,1)) +
  guides(col = F) +
  coord_fixed(ratio = 1) +
  theme_bw()
```

## Country-wide trend
Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics[^3]. The baseline of party preference has varied over the elections, with the biggest swing occuring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.

[^3]: Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.

```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.", out.height="15%", out.extra = ''}
intercept_df <- bind_rows(data.frame(summary(glsmod16)$tTable)[1, ]  %>% mutate(year = "2016"), 
  data.frame(summary(glsmod13)$tTable)[1, ] %>% mutate(year = "2013"),
  data.frame(summary(glsmod10)$tTable)[1, ] %>% mutate(year = "2010"), 
  data.frame(summary(glsmod07)$tTable)[1, ] %>% mutate(year = "2007"),
  data.frame(summary(glsmod04)$tTable)[1, ] %>% mutate(year = "2004"), 
  data.frame(summary(glsmod01)$tTable)[1, ] %>% mutate(year = "2001")) %>% 
  rename(se = `Std.Error`, estimate = Value, p = `p.value`) %>% 
  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se)

intercept_df %>% ggplot() + 
  geom_point(aes(x = year, y = estimate), size = 3) +
  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) + 
  geom_hline(aes(yintercept = 50), alpha = 0.5, size = 1) +
  labs(x = "Election year", y = "Estimate of intercept") +
  guides(col = F) +
  coord_fixed(ratio = 0.2) +
  lims(y = c(45, 55)) +
  theme_bw()
```

## Influential socio-demographics
To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used. These show the direction, size and significance of an estimated effect - the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band. If a horizontal line can be drawn through the confidence band, then the effect is insignificant. Plots for each election are faceted to compare the effects over time in Figure \@ref(fig:partresplot1) and Figure \@ref(fig:partresplot2).

It is important here to note the ecological fallacy - insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).

```{r myvisreg, include = F}
# Function to produce visreg style conditional plots
my_visreg <- function(my_model, sp_weights, varname, 
  plot = FALSE, nolabs = FALSE, xlimits = NULL, ylimits = NULL, year = "") {

@@

  if (!is.null(xlimits) & !is.null(ylimits)) {
    myplot <- myplot + coord_cartesian(xlim = xlimits, ylim = ylimits)
  }
      return(myplot)
  }

  # Points
  if (plot == FALSE) {
    return_ls <- list(bands = plot_df, points = points_df)

    return(return_ls)
  }

# Grid visreg
grid_visreg <- function(varname, plot = TRUE, myscale = "free", top = FALSE) {

    p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, plot = F, year = "2016")
    p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, plot = F, year = "2013")
    p10 <- my_visreg(glsmod10, sp_weights_10, varname = varname, plot = F, year = "2010")
    p07 <- my_visreg(glsmod07, sp_weights_07, varname = varname, plot = F, year = "2007")
    p04 <- my_visreg(glsmod04, sp_weights_04, varname = varname, plot = F, year = "2004")
    p01 <- my_visreg(glsmod01, sp_weights_01, varname = varname, plot = F, year = "2001")
